---
title: "Marketing Analytics Project 1 - Conjoint Analysis (Survey Analytics)"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

### Reading the Fashion Data

```{r}
#Loading Data
data <- read.delim("/Users/Monu/Desktop/Marketing Analytics/A2/fashiondata.txt", header = TRUE, sep = " ", dec = ",")

#Creating Person ID 
b<-c()
for (i in 1:400){
  a <- rep(i,8)
  b <- c(b,a) 
}
data$Person_ID <- b

#Viewing Data
data[1:16,]
```

Setting Up the design matrix (effects coded).

```{r}
#Creating Design Mat
dat_des <- data[,-5:-6]
dat_des <- dat_des[1:8,2:4]
dat_des

#Creating Effects Coding
des.mm <- model.matrix(~FASHION+QUALITY+PRICE,data=dat_des,contrasts.arg=list(FASHION=contr.sum,QUALITY=contr.sum,PRICE=contr.sum))
des.mm <- des.mm[,2:4]
des.mm
cor(des.mm)
```

Data Frame with all data and effects-coded variables. 

```{r}
data_ef <- cbind(data[data$Person_ID==1,c(1,7)],des.mm)

for(i in 2:400){
pp_temp <- cbind(data[data$Person_ID==i,c(1,7)],des.mm)
data_ef <- rbind(data_ef,pp_temp)
}

data_ef <- data.frame(data_ef)
names(data_ef) <- c( "RATING","PERSON_ID","FASHION1","QUALITY1","PRICE1")

data_ef[1:16,]
dim(data_ef)
```

## 1. Parthworth estimation for all individuals separately.

```{r}

####Initializing the Data for First Person 
##Train and Test
trainIND_per_person_ <- sample(seq_len(8), size = 6)
train_pp <- data_ef[data_ef$PERSON_ID==1,][trainIND_per_person_,]
test_pp <- data_ef[data_ef$PERSON_ID==1,][-trainIND_per_person_,]

##Individual Model Per Person 
model_pp <- lm(RATING~FASHION1+QUALITY1+PRICE1,data=train_pp)
model_data <- data.frame(model_pp$coefficients)

#Predicting Train and Test
train_pred_pp <- predict(model_pp)
test_pred_pp <- predict(model_pp,test_pp)

all_data <- rbind(cbind(train_pp,indv.pred=train_pred_pp,split=c(rep('train'))),cbind(test_pp,indv.pred=test_pred_pp,split=c(rep('test'))))

####For Loop for Each Person after First Person 
##Train and Test
for(i in 2:400){
trainIND_per_person_ <- sample(seq_len(8), size = 6)
train_pp <- data_ef[data_ef$PERSON_ID==i,][trainIND_per_person_,]
test_pp <- data_ef[data_ef$PERSON_ID==i,][-trainIND_per_person_,]

##Individual Model Per Person 
model_pp <- lm(RATING~FASHION1+QUALITY1+PRICE1,data=train_pp)
coef_pp <- data.frame(model_pp$coefficients)

#Predicting Train and Test
train_pred_pp <- predict(model_pp)
test_pred_pp <- predict(model_pp,test_pp)

#Creating Per Person Data
per_person <- rbind(cbind(train_pp,indv.pred=train_pred_pp,split=c(rep('train'))),cbind(test_pp,indv.pred=test_pred_pp,split=c(rep('test'))))

#Appending to All Data
all_data <- rbind(all_data,per_person)
model_data <- cbind(model_data,coef_pp)
}
```


Data Frame with Coeffcients from all 400 Individual Models. 

```{r}
model_data[,1:5]
dim(model_data)
```

Data Frame with all data including predictions for test and train profiles.

```{r}
all_data[1:16,]
dim(all_data)
```

Evaluating Model Performance

```{r}
ind_model_perf <-  cbind(train_performance=(cor(all_data[all_data$split=='train',]$RATING,all_data[all_data$split=='train',]$indv.pred))^2,test_performance=(cor(all_data[all_data$split=='test',]$RATING,all_data[all_data$split=='test',]$indv.pred))^2)

ind_model_perf
```

## 2. Parthworth estimation by combining all individual responses into an aggregate. 

```{r}
agg_model_train <- all_data[all_data$split=='train',c('RATING','FASHION1','QUALITY1','PRICE1')]
dim(agg_model_train)

agg_model_test <- all_data[all_data$split=='test',c('RATING','FASHION1','QUALITY1','PRICE1')]
dim(agg_model_test)

#Fitting the Model 
agg_model <- lm(RATING~FASHION1+QUALITY1+PRICE1,data=agg_model_train)
summary(agg_model)

#Predicting on Train and Test Set
agg_pred_train <- predict(agg_model)
agg_pred_test <- predict(agg_model,agg_model_test)
```

Evaluating Model Performance

```{r}
agg_model_perf <- cbind(
train_performance = (cor(agg_model_train$RATING,agg_pred_train))^2,
test_performance = (cor(agg_model_test$RATING,agg_pred_test))^2
)

agg_model_perf
```

## 3. Parthworth estimation by using linear mixed models.
In this case you should treat each individual as a seperate group for defining the random effects.

```{r}
require(lme4)

#Train and Test Datasets
mixed_model_train <- all_data[all_data$split=='train',c('RATING','FASHION1','QUALITY1','PRICE1','PERSON_ID')]
dim(mixed_model_train)

mixed_model_test <- all_data[all_data$split=='test',c('RATING','FASHION1','QUALITY1','PRICE1','PERSON_ID')]
dim(mixed_model_test)

#Fitting Mixed Model using lmer
mixed_model <- lmer(RATING~FASHION1+QUALITY1+PRICE1+(1+FASHION1+QUALITY1+PRICE1|PERSON_ID),  control=lmerControl(optCtrl = list(maxfun=100000)), data=mixed_model_train)
summary(mixed_model)

#Predicting on Train and Test Set
mixed_pred_train <- predict(mixed_model)
mixed_pred_test <- predict(mixed_model,mixed_model_test)
```

```{r}
#Fixed Effects
fixef(mixed_model)

#Random Effects for each Individual 
head(ranef(mixed_model)$PERSON_ID)

#Final Coeffcieints (Fixed + Random Effects)
head(coef(mixed_model)$PERSON_ID)
```

Evaluating Model Performance 
```{r}
mixed_model_perf <- cbind(
train_performance = (cor(mixed_model_train$RATING,mixed_pred_train))^2,
test_performance = (cor(mixed_model_test$RATING,mixed_pred_test))^2
)

mixed_model_perf
```

## 4. Partworth estimation of the mixed effects models by using MCMChregress 

```{r}
require(MCMCpack)

set.seed(1)
mc_model=MCMChregress(fixed=RATING~FASHION1+QUALITY1+PRICE1,
                      random =~1+FASHION1+QUALITY1+PRICE1, 
                      group="PERSON_ID", 
                      r=4,
                      R=4*diag(4), 
                      data=mixed_model_train)
```

```{r}
#summary(mc_model$mcmc)
dim(mc_model$mcmc)
```


```{r}
##Getting MCMC Model Fixed and Random Coefs Means

#Taking only the last 500 rows to avoid taking the mean from the samples in the first half of MCMC process 
mc_coef_samples <- mc_model$mcmc[501:1000,1:1604]

#Getting Mean of Each Sample 
mc_coef_samples_mean <- apply(mc_coef_samples,2,mean)

#Fixed Coef Mean and Random Coef Means
fixed_mean_values <- mc_coef_samples_mean[1:4]
random_mean_values <- mc_coef_samples_mean[5:1604]

#intercept coefcieints per individual
int_coef <- random_mean_values[1:400]
int_coef <- int_coef + fixed_mean_values[1]
names(int_coef) <- substr(names(random_mean_values[1:400]),15,18)

#fashion coefcieints per individual
fash_coef <- random_mean_values[401:800]
fash_coef <- fash_coef + fixed_mean_values[2]
names(fash_coef) <- substr(names(random_mean_values[1:400]),15,18)

#quality coefcieints per individual
qual_coef <- random_mean_values[801:1200]
qual_coef <- qual_coef + fixed_mean_values[3]
names(qual_coef) <- substr(names(random_mean_values[1:400]),15,18)

#price coefcieints per individual
price_coef <- random_mean_values[1201:1600]
price_coef <- price_coef + fixed_mean_values[4]
names(price_coef) <- substr(names(random_mean_values[1:400]),15,18)
```


```{r}
#Manually Calculating MCMC Predicted Values on Test Set

#initializing for person 1
#formula for calculating the predicted value
value <- int_coef[names(int_coef)==1]+mixed_model_test[mixed_model_test$PERSON_ID==1,'FASHION1']*fash_coef[names(fash_coef)==1]+mixed_model_test[mixed_model_test$PERSON_ID==1,'QUALITY1']*qual_coef[names(qual_coef)==1]+mixed_model_test[mixed_model_test$PERSON_ID==1,'PRICE1']*price_coef[names(price_coef)==1]

mcmc_test_pred <- cbind(rating=mixed_model_test[mixed_model_test$PERSON_ID==1,'RATING'],mcmc.pred=value)

#for loop for rest of the individuals after the first person 
for(i in 2:400){
value_pp <- int_coef[names(int_coef)==i]+mixed_model_test[mixed_model_test$PERSON_ID==i,'FASHION1']*fash_coef[names(fash_coef)==i]+mixed_model_test[mixed_model_test$PERSON_ID==i,'QUALITY1']*qual_coef[names(qual_coef)==i]+mixed_model_test[mixed_model_test$PERSON_ID==i,'PRICE1']*price_coef[names(price_coef)==i]

temp_pp <- cbind(rating=mixed_model_test[mixed_model_test$PERSON_ID==i,'RATING'],mcmc.pred=value_pp)

mcmc_test_pred <- rbind(mcmc_test_pred,temp_pp)
}
```

```{r}
head(mcmc_test_pred)
dim(mcmc_test_pred)
```

Evaluating Model Performance

```{r}
mcmc_model_perf<-cbind(
train_performance=cor(mixed_model_train$RATING,mc_model$Y.pred)^2,
test_performance=cor(mcmc_test_pred[,1],mcmc_test_pred[,2])^2)

mcmc_model_perf
```


## 5. Assess performance of 1-4 using 2 profiles as "Holdout" profiles for each 

Below the test performance shows how the hold-out profiles did for each type of model.

```{r}
model_perf_summary <- rbind(
ind_model_perf,
agg_model_perf,
mixed_model_perf,
mcmc_model_perf
)

rownames(model_perf_summary) <- c("Individual Models R-Squared","Aggregate Models R-Squared","Mixed Models R-Squared","MCMC Models R-Squared")

model_perf_summary
```

The Mixed Model had the highest R-Squared on the hold-out (test) set. Mixed Model will be used for clustering of the individual's parthworths in step 6. 

# 6. Applying K-Means Clustering to the Partworths from the Best Performing Model

```{r}
#Loading all Clustering Packages
require('fpc')
require('factoextra')
require('cluster')
require('FunCluster')
```


```{r}
#Getting the coefficients from Mixed Model 
mixed_model_coef <- coef(mixed_model)$PERSON_ID

#Scaling the coefficient values
mixed_model_coef <- scale(mixed_model_coef)
```

Determining what the optimal number of cluster should be with the elbow method.

```{r}
fviz_nbclust(mixed_model_coef,FUNcluster = kmeans,method='wss')
```

Applying t-SNE to further visualize the clusters

```{r}
require('Rtsne')
tsne_model <- Rtsne(as.matrix(mixed_model_coef), check_duplicates=FALSE, pca=TRUE, perplexity=100, theta=0.5, dims=2)
d_tsne_1 = as.data.frame(tsne_model$Y)  
```

```{r}
## Plotting the Results without Clustering
ggplot(d_tsne_1, aes(x=V1, y=V2)) +  
  geom_point(size=0.25) +
  guides(colour=guide_legend(override.aes=list(size=6))) +
  xlab("") + ylab("") +
  ggtitle("t-SNE") +
  theme_light(base_size=20) +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank()) +
  scale_colour_brewer(palette = "Set2")
```

Based on these findings, 3 clusters will be fit to the model. 

#### Fitting K-Means with 3 Clusters

```{r}
clust <- kmeans(mixed_model_coef,3,iter.max = 20)
```

Visualizing the Clusters

```{r}
plotcluster(mixed_model_coef, clust$cluster)
```

According to the data, these three segments can be used as the benefit segments for the fashion product. 

```{r}
require(dplyr)
row_ind <- seq(1,3200,8)

cluster_mat <- cbind(coef(mixed_model)$PERSON_ID,clusters = clust$cluster,data[row_ind,c('GENDER','AGE')])

names(cluster_mat) <- c('int','fash','qual','price','clusters','gender','age')

head(cluster_mat)
```


```{r}
#Mean Values of Coeffcients for each Cluster
cluster_mat %>% group_by(clusters) %>% summarise(int = mean(int),
                                                 fash=mean(fash),
                                                 qual=mean(qual),
                                                 price=mean(price))

```

Fashion: 
Clusters 2 and 3 prefer modern fashion (higher change in preference from Traditional to Modern) more than cluster 1. Cluster 3 prefers modern fashion the most. 

Quality: 
Cluster 1 is the most sensitive to high quality, and has the highest mean coefficient for high quality. Clusters 2 and 3 are less sensitive about high quality. Cluster 3 has the lowest quality mean coefficient indicating that there is no additional perceived value of high quality for cluster 3. 

Price:
Cluster 3 is slightly more sensitive to higher prices than clusters 1 and 2. 


```{r}
cluster_mat %>% group_by(clusters,age) %>% summarise(n=n())
```

Age Breakdown: 
Cluster 1 - mostly 40+ individuals 

Cluster 2 - mostly younger (16-39) individuals 

Cluster 3 - primarily younger individuals with 16-24 

```{r}
cluster_mat %>% group_by(clusters,gender) %>% summarise(n=n())
```

Gender Breakdown: 

Cluster 1 and 2 - Mix of Male and Female 

Cluster 3 - Mostly Female 

*In summary, here are the three broad segments:* 

Segment 1 - Older Individuals with High Quality Purchases - Care the most about high quality and prefer modern fashion the least (traditional fashion and modern fashion don't have high preference differences). The majority of them are 40+. 

Segment 2 - Modern, Fashion-Forward Young Professionals - Prefer Modern Fashion and care less about high quality - they are also mostly younger individuals (16-24) and some older professionals (24 - 39 or 40+). 

Segment 3 - Students / Younger Indiviudals [Mostly Female] - Most price sensitive, prefer modern fashion to  to traditional fashion the most, and care the least about high quality. The majority of them are younger individuals (16-24) and females. 
